Mathematics of Reinforcement Learning with Applications to Quantitative Finance

### 2022-08-23 papers/paper_2-col.tex
add reference and re-format paper

% !TEX encoding = UTF-8 Unicode
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%
%%  This file is asmeconf-template.tex, a LaTeX template to format ASME Conference papers according to
%%  the requirements on ASME's conference web pages, and including hypertext support for the pdf.
%%
%%  This file is version 1.31 dated 2022/07/04
%%  
%%  As of version 1.11, this template defaults to ASME's newer conference guidelines first posted July 2019.
%% 			Those guidelines changed the requested author block formatting to be inline. 
%%			This LaTeX template continues to support the traditional grid format as a package option, [grid].
%%			Nomenclature now follows the abstract. Abstract text is set in italics.
%%
%%  Author: John H. Lienhard V
%%          Department of Mechanical Engineering
%%          Massachusetts Institute of Technology
%%          Cambridge, MA 02139-4307 USA
%%
%%  Class options include:
%%
%%          * Math options from M. Sharpe's newtxmath package: upright integrals [upint]; 
%%          *    [varvw] for a v and w that are better distinguished from Greek nu; and also 
%%          *    [subscriptcorrection, smallerops, varg, frenchmath, varbb, cmbraces, slantedGreek,...] 
%%			* 	 See newtx documentation for descriptions (at CTAN: http://ctan.org, v1.6 or higher is best).
%%
%%          * Options to: 	omit the ASME copyright footer [nofoot]; 
%%			*				use government employee copyright notice [govt];
%%			*				use government contractor copyright notice [contractor]
%%
%%          * An option to use the traditional grid arrangement of author names [grid].  
%%			*	 With this option, line breaks (\\) may be inserted into the address as needed. 
%%			*	 Author names that include commas should be enclosed in braces, e.g., {Joseph L. Smith, Jr.}.  
%%			*	 Authors may be grouped above a single affiliation using braces, e.g., {Henry Tudor, Catherine Parr}.
%%
%%          * An option to balance the heights of columns on the last page [balance]. 
%%          *    This option is NOT compatible with the [lineno] option.
%%
%%          * An option to include line numbers [lineno]. You must *run twice* for proper placement of the line numbers. 
%%			*	 The lineno package does not number titles, footnotes, captions, or tables.
%%          *    This option will disable balancing column height on final page if that option has been invoked.
%%          *    The lineno package won't always number the lines preceding displayed math in a paragraph because
%%          *    paragraph has not ended.  See that package's documentation for macros to address this problem, or
%%          *    just leave a blank line above the displayed equation while you are editing and then remove the 
%%          *    blank line and [lineno] option when you move to your final version.
%%
%%			* Options for PDF/A compliance. [pdf-a] will produce PDF/A-3u compliance with sRGB OutputIntent.
%%			*	 [pdfapart= 1 or 2 or 3] and [pdfaconformance= b or u] can enable levels 1b, 2b, 2u, and 3b.
%%			*    The most recent versions of LaTeX (2021 and later) are moving toward integrated support for pdf-a, 
%%			*    through \DeclareDocumentMetadata{..}. The asmeconf class supports these new features, which can 
%%			*	 replace the aforementioned class options. (An up-to-date LaTeX installation is required to use this.)
%%
%%          * Many options for calligraphic, script, and fraktur fonts from the mathalfa package; the
%%          *    example shown here is: [mathalfa=cal=boondoxo] to use a Boondox font for \mathcal.
%%          *    Some other options for cal are: dutchcal, zapfc, cm (default), euler,...
%%          *    frak (fraktur), bb (blackboard bold), scr (script) may also be chosen this way.
%%			*	 For details, refer to mathalfa documentation (at CTAN: http://ctan.org).
%%
%%          * Option to use superiors font from newtxtext for footnotes [nodefaultsups] and
%%          *    for slightly larger small capitals, [largesc], also from newtxtext.
%%
%%          * An option to allow hyphenation of the typewriter font [hyphenate], from inconsolata package.
%%          *    Hyphenation is normally suppressed for typewriter mode because it is often used for code.
%%			*	 To replace the default variable word spacing by monospacing, use the option [mono].
%%			*	 To get a zero without a slash, use [var0]
%%
%%          * Options (used by the babel package) to include passages in languages other than English (e.g., a translation 
%%			*    of the abstract). Languages are called as options, e.g. [french], [spanish], [greek], [russian], etc. 
%%			*    Language support is greatest when running LuaLaTeX with the [fontspec] option.
%%			*    See Appendix B for details.
%%
%%  The use of commands defined or modified by the asmeconf class is illustrated throughout this file. In particular, 
%%  ASME requires capitalized, sans-serif section headings, and as a result some care is needed when using macros 
%%  in section headings, as also illustrated below.
%%
%%  Use an up-to-date and complete installation of LaTeX, such as TeX Live 2020 or later. Errors may occur with old set-ups.
%%
 %=========================================================
%% 
%% LICENSE: 
%%
%% Copyright (c) 2022 John H. Lienhard
%%
%% Offered under the MIT license: https://ctan.org/license/mit 
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Class options are described above. Change these options as desired. 
%%		If you are not using the language options, remove them (together with Appendices B and C)
%%	 	Remove the [colorlinks] option before *final* submission to ASME, to get black text for printing,
%%		but keep that option for other uses.

\documentclass[balance,upint,subscriptcorrection,varvw,mathalfa=cal=boondoxo,pdf-a,colorlinks,nofoot]{asmeconf}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta}
\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=1.3cm, text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>={Stealth[scale=1.1]}]
\usepackage{algorithm, algpseudocode}
\usepackage{siunitx}
\graphicspath{ {../data/} }
%%%%%  pdf metadata  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypersetup{%
	pdfauthor={Yuanhao JIANG},
	pdftitle={Mathematics of Reinforcement Learning with Applications to Quantitative Finance},
	pdfkeywords={Reinforcement Learning,  Artificial Neural Networks},
	pdfsubject={Describe Mathematics of Reinforcement Learning with Applications to Quantitative Finance},
    pdfurl={https://github.com/Yuanhao-JIANG/RL-in-QF},
%	pdflicenseurl={https://ctan.org/pkg/asmeconf},
}

\begin{document}
% Change these fields to the right content for your conference.
% You can comment these out if for some reason you don't want a header.
% Use title case (first letters capitalized), not all capitals
\ConfName{Proceedings of the ASME 2022\linebreak International Mechanical Engineering Congress and Exposition}
\ConfAcronym{IMECE2022}
\ConfDate{October 30--November 3, 2022} % update 
\ConfCity{Columbus, OH} % update 
\PaperNo{IMECE2022-XXXX}

% Units of measure (e.g., cm) and other specialty lowercase terms in the title should be 
%   enclosed in \NoCaseChange{...} to maintain lower case type
%   LaTeX will automatically set the rest of the title in all capital letters.

\title{Mathematics of Reinforcement Learning with Applications to Quantitative Finance}
%\title{Place Title Here: Place Subtitle After Colon}

%   Put author names into the order you want. Use the same order for affiliations.
%   \affil{#} tags the author's affiliation to the address in \SetAffiliation{#}.
%   No space between last name and \affil{#}, separate names with commas.
%
%	For a sole author or a single affiliation for all authors, {#} may be left empty, as \affil{} and \SetAffiliation{} (but not with [grid] option!)
%
%   \CorrespondingAuthor{email} follows that author's affiliation, no spaces.  
%   If multiple corresponding authors, put both email addresses in the same command and place after both authors.
%
%   \JointFirstAuthor, if applicable, follows the affiliation of the relevant authors, no spaces.
\SetAuthors{
	Yuanhao JIANG\affil{1}\CorrespondingAuthor{s2132254@ed.ac.uk}
}

\SetAffiliation{1}{The University of Edinburgh, Edinburgh, UK}
%	To switch from inline author names to gridded names, use the [grid] option.

@@

%%% Change these to your keywords.  Keywords are automatically printed at the end of the abstract.
%%% This command MUST COME BEFORE the end of the abstract.
%%% If you don't want keywords, leave the argument of \keywords{} empty (or use the abstract* environment)
\keywords{
    Reinforcement Learning, Artificial Neural Networks, Quantitative Finance
}

%% Abstract should be 200 words or less
\begin{abstract}
\noindent This paper provides a new way of pricing in the car insurance
industry, which is, using reinforcement learning algorithms to train pricing 
agents to promote proper price, such that our customer is most likely to 
purchase our product, and we also get the highest profit and a better 
portfolio. 
We first introduce general reinforcement learning knowledge, including how 
we model the interaction between agent and environment using Markov decision
process, and the mathematics behind those methods to solve the reinforcement 
learning problem, especially the Actor-Critic algorithm.
Then we start building the environment, and use artificial neural networks as 
our pricing models, and writing algorithms that are specific to our financial 
problem. 
In the end, we will compare the Actor-Critic algorithms with other popular 
algorithms including REINFORCE and PPO in terms of their rate of convergence 
to a stable reward.
\end{abstract}

@@

        \draw [arrow] (env.163) -- ++(-1.5,0) node[anchor=south, xshift=0.8cm]
            {\(S_{t+1}\)} coordinate(a1);
        \draw [arrow] (a1) -- ++(-0.7,0) |- node[anchor=south, xshift=1.3cm] 
            {\(S_t\)} (pi);
        \draw [arrow] (env.197) -- ++(-1.5,0) node[anchor=south, xshift=0.8cm]
            {\(R_{t+1}\)} coordinate(a2);
        \draw [arrow] (a2) -- ++(-1.3,0) -- ++(0,3) 
            -- node[anchor=south, xshift=0.5cm] {\(R_t\)} (agent.159);
        \draw [arrow] (pi) -- ++(2.7,0) node[anchor=south, xshift=-0.8cm] 
            {\(A_t\)} |- (env);
        \draw [dashed,thick] (-2.7,-1.5) -- (-2.7, -3.2);
    \end{tikzpicture}
    \caption{MDP Illustration}
    \label{fig:MDP-illustration}
\end{figure}
\noindent However, in RL, the agent is not told which actions to take, but 
instead must discover which actions yield the most reward by trying them. Our 
goal at each time step is to find a policy \(\pi\) to maximize the expected 
reward alonge the process afterwards, which is given by the value function
\[V(s)=\mathbb{E}\left[\sum_{i=0}^{\infty} \gamma^i R_{t+i+1}\;\middle|\; 
S_t=s\right]\]
To achieve this, we need to evaluate the current policy, and improve it 
accordingly.

@@

\subsection{Policy Iteration}
\noindent By iteratively applying policy evaluation and policy improvement, we
can obtain a sequence of monotonically improving policies and value functions:
\[\pi_0 \overset{E}{\longrightarrow} v_{\pi_0} \overset{I}{\longrightarrow}
\pi_1 \overset{E}{\longrightarrow} v_{\pi_1} \overset{I}{\longrightarrow}
\dots \overset{I}{\longrightarrow} \pi_* \overset{E}{\longrightarrow} v_*\]
where \(\overset{E}{\longrightarrow}\) denotes a policy evaluation and
\(\overset{I}{\longrightarrow}\) denotes a policy improvement. This way of 
finding an optimal policy is called policy iteration.

\subsection{Value Iteration}
\noindent One drawback to policy iteration is that each of its iterations 
involves policy evaluation, which itself is a protracted iterative computation
requiring multiple sweeps through the state set.
\vspace{3mm}\newline We now truncate the policy evaluation such that it is
stopped after just one sweep (one update of each state):
\begin{align*}
    v_{k+1}(s)&=\underset{a}{\text{max}}\:q(s, a)\\
              &=\underset{a}{\text{max}}\:\mathbb{E}\left[
          R_{t+1}+\gamma v_k(S_{t+1})\;\middle|\;S_t=s, A_t=a\right]
\end{align*}
We keep iterate until \(v\) converges (to \(v_*\)), then we record the last
policy \(\pi(s) = \underset{a}{\text{argmax}}\:q(s, a)\).

\subsection{Generalized Policy Iteration}
\noindent As long as both processes, policy evaluation and policy improvement,
continue to update all states, the ultimate result is typically the 
same --- converge to the optimal value function and an optimal policy.
\vspace{3mm}\newline Generalized policy iteration (GPI) is used to refer to the
general idea of letting policy evaluation and policy improvement interact,
independent of the granularity of the two processes (instead of letting each
complete before the other begins).

@@

\subsection{Policy Gradient Methods}
\noindent When the state space is arbitrarily large, we cannot find an optimal 
policy or the optimal value function given the limited resources and time. 
Instead of keeping track of the value and the action(s) to select for each 
state (tabular methods), we can use parameterized functions, for example, 
artificial neural networks (ANNs), to approximate the value function and the 
policy function.
\vspace{3mm}\newline Previously methods learned the values of actions and then
select actions based on their estimated action values (q function). Now we
consider methods that instead learn a parameterized policy that select actions
without consulting a value function. A value function may still be used to
learn the policy parameter, but is not required for action selection.
\vspace{3mm}\newline We use \(\bm{\theta}\) and \(\bm{w}\) to denotes the
policy's parameter vector and value function's weight vector, respectively.
Then we have our policy \(\pi\left(a\;\middle|\;s, \bm{\theta}\right)
=\text{Pr}\left\{A_t=a\;\middle|\;S_t=s, \bm{\theta}_t=\bm{\theta}\right\}\) and
value function \(v_\pi(s, \bm{w})\). We also define the scalar performance
measure (objective) to be
\[J(\bm{\theta})=\begin{cases}
    \ v_{\pi_{\bm{\theta}}}(s_0); &\text{episodic case}\\
    \ \lim_{t \to \infty} \mathbb{E}
    \left[R_t\;\middle|\;S_0,A_{0:t-1}\sim\pi\right];
                                  & \text{continuing case}
\end{cases}\]
Policy gradient methods seeks to maximize this performance measure, so their
updates approximate gradient ascent in \(J\):
\[\bm{\theta}_{t+1}=\bm{\theta}_t+\alpha\widehat{\nabla J(\bm{\theta}_t)}\]
where \(\widehat{\nabla J(\bm{\theta}_t)}\) is a stochastic estimate whose
expection approximates the gradient of the performance measure w.r.t
\(\bm{\theta}_t\).

@@

\subsection{Training Result}
\noindent Figure~\ref{fig:a2c-out} shows the training result of 
Algorithm~\ref{algorithm:actor-critic} (300 moving average reward vs 
iteration), with 600 iterations (5 episodes per iteration), \(T=300\), 
\(\gamma=0.99\), \(\alpha_\theta=\alpha_{\bm{w}}=\num{3e-4}\).
\begin{figure*}
    \centering\includegraphics[scale=0.75]{a2c_out}
    \caption{Training Result for Algorithm a2c}
    \label{fig:a2c-out}
\end{figure*}
The graph shows that the moving average reward converges within only 50
iterations, that is, 250 episode, 75000 parameter updates for each network,
and stays quite stable afterwards.

\section{Actor-Critic, Reinforce, and PPO}
\subsection{Algorithms}
\noindent To further show how Actor-Critic algorithm performs, we compare it 
to another two popular algorithms: Reinforce 
(Algorithm~\ref{algorithm:reinforce}) and PPO (Algorithm~\ref{algorithm:ppo}).
The policy function of the two algorithms is implemented the same way as we 
did in Actor-Critic algorithm, and the same is true for the value function 
used in PPO method.
\begin{algorithm}
    \caption{Reinforce for financial problem}\label{algorithm:reinforce}
    \begin{algorithmic}[1]
        \State Initialize ANN policy parameterization
        \(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
        \State Step sizes \(\alpha_{\bm{\theta}}>0\)
        \Loop \hspace{0.3mm} (for each episode)
        \State Reset the environment
        \State Generate an episode 
        \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
        following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
        \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
        \EndFor
        \EndLoop
    \end{algorithmic}
\end{algorithm}

\begin{algorithm*}
    \caption{PPO for financial problem}\label{algorithm:ppo}
    \begin{algorithmic}[1]
        \State Initialize ANN policy parameterization
        \(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
        \State Initialize ANN state-value function parameterization
        \(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
        \State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
        \For{iteration \(k=0, 1, 2, ...\)}
        \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\),
        following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}_k\right)\), where
        \(\tau = S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\)
        \For{\(t=0, 1, ..., T\) in each trajectory \(\tau\) in 
        \(\mathcal{D}_k\)}
        \State \(G_t \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(A_t \gets G_t - \hat{v}(S_t)\)
        \EndFor
        \Loop \hspace{0.3mm} (update multiple times)
        \State \(\bm{\theta} \gets \underset{\bm{\theta}}{\text{arg\:max}}
        \frac{1}{|\mathcal{D}_k|T}
        \sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T}
        \min{\left(
                \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
                {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)}A_t,
                \text{clip}\left(
                    \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
                    {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)},
                    1-\epsilon, 1+\epsilon
                \right)A_t
        \right)}\)
        \State \(\bm{w} \gets \underset{\bm{w}}{\text{arg\:min}}
        \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T}
        \Bigl(\hat{v}(S_t, \bm{w})-G_t\Bigr)^2\)
        \EndLoop
        \EndFor
    \end{algorithmic}
\end{algorithm*}

\subsection{Training result}
\noindent Figure~\ref{fig:3-algorithms-out} shows the training 
result of these three algorithms (300 moving average reward vs iteration), 
with 600 iterations (5 episodes per iteration), \(T=300\), \(\gamma=0.99\),
\(\alpha_\theta=\num{3e-4}\), \(\alpha_{\bm{w}}=\num{3e-10}\) (if applicable),
\(\epsilon=0.5\) (if applicable).
\begin{figure*}
    \centering\includegraphics[scale=0.75]{out}
    \caption{Training Result Comparison for Three Algorithm}
    \label{fig:3-algorithms-out}
\end{figure*}
As we can see, the Actor-Critic algorithm converges the fastest, within 50 
iterations, followed by REINFORCE algorithm, which converges within 250 
iterations, despite a small drop of moving average around iteration 300, 
which is quickly corrected by the algorithm within 50 iterations. PPO 
algorithm converges the slowest, with about 300 iterations.

@@

\section*{Acknowledgments}
\noindent Place any acknowledgments here.

>>>>

\documentclass[balance,upint,subscriptcorrection,varvw,mathalfa=cal=boondoxo,pdf-a,nofoot]{asmeconf}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta}
\tikzstyle{process} = [
    rectangle, minimum width=2cm, minimum height=1.3cm, text centered, 
    draw=black
]
\tikzstyle{arrow} = [thick,->,>={Stealth[scale=1.1]}]
\usepackage{algorithm, algpseudocode}
\usepackage{siunitx}
\graphicspath{ {../data/} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  pdf metadata  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\hypersetup{
	pdfauthor={Yuanhao JIANG},
	pdftitle={
        Mathematics of Reinforcement Learning with Applications to 
        Quantitative Finance
    },
	pdfkeywords={
        Reinforcement Learning,  Artificial Neural Networks, Quantitative 
        Finance
    },
	pdfsubject={
        Describe Mathematics of Reinforcement Learning with Applications to 
        Quantitative Finance
    },
    pdfurl={https://github.com/Yuanhao-JIANG/RL-in-QF},
}

\begin{document}
\title{
    Mathematics of Reinforcement Learning with Applications to Quantitative 
    Finance
}

\SetAuthors{
	Yuanhao JIANG\affil{1}\CorrespondingAuthor{s2132254@ed.ac.uk}
}

\SetAffiliation{1}{The University of Edinburgh, Edinburgh, UK}

@@

\keywords{
    Reinforcement Learning, Artificial Neural Networks, Quantitative Finance
}

\begin{abstract}
\noindent This paper provides a new way of pricing in the car insurance
industry, which is, using reinforcement learning algorithms to train pricing 
agents to promote proper price, such that our customer is most likely to 
purchase our product, and we also get the highest profit and a better 
portfolio. 
We first introduce general reinforcement learning knowledge, including how 
we model the interaction between agent and environment using Markov decision
process, and the mathematics behind those methods to solve the reinforcement 
learning problem, especially the Actor-Critic algorithm.
Then we start building the environment, and use artificial neural networks as 
our pricing models, and writing algorithms that are specific to our financial 
problem. 
In the end, we will compare the Actor-Critic algorithms with other popular 
algorithms including REINFORCE and PPO in terms of their rate of convergence 
to a stable reward.
\end{abstract}

@@

        \draw [arrow] (env.163) -- ++(-1.5,0) node[anchor=south, xshift=0.8cm]
            {\(S_{t+1}\)} coordinate(a1);
        \draw [arrow] (a1) -- ++(-0.7,0) |- node[anchor=south, xshift=1.3cm] 
            {\(S_t\)} (pi);
        \draw [arrow] (env.197) -- ++(-1.5,0) node[anchor=south, xshift=0.8cm]
            {\(R_{t+1}\)} coordinate(a2);
        \draw [arrow] (a2) -- ++(-1.3,0) -- ++(0,3) 
            -- node[anchor=south, xshift=0.5cm] {\(R_t\)} (agent.159);
        \draw [arrow] (pi) -- ++(2.7,0) node[anchor=south, xshift=-0.8cm] 
            {\(A_t\)} |- (env);
        \draw [dashed,thick] (-2.7,-1.5) -- (-2.7, -3.2);
    \end{tikzpicture}
    \caption{MDP Illustration~\cite{Sutton1998}}
    \label{fig:MDP-illustration}
\end{figure}
\noindent However, in RL, the agent is not told which actions to take, but 
instead must discover which actions yield the most reward by trying them. Our 
goal at each time step is to find a policy \(\pi\) to maximize the expected 
reward alonge the process afterwards, which is given by the value function
\[V(s)=\mathbb{E}\left[\sum_{i=0}^{\infty} \gamma^i R_{t+i+1}\;\middle|\; 
S_t=s\right]\]
To achieve this, we need to evaluate the current policy, and improve it 
accordingly.

@@

\subsection{Policy Iteration}
\noindent By iteratively applying policy evaluation and policy improvement, we
can obtain a sequence of monotonically improving policies and value 
functions~\cite{Sutton1998} :
\[\pi_0 \overset{E}{\longrightarrow} v_{\pi_0} \overset{I}{\longrightarrow}
\pi_1 \overset{E}{\longrightarrow} v_{\pi_1} \overset{I}{\longrightarrow}
\dots \overset{I}{\longrightarrow} \pi_* \overset{E}{\longrightarrow} v_*\]
where \(\overset{E}{\longrightarrow}\) denotes a policy evaluation and
\(\overset{I}{\longrightarrow}\) denotes a policy improvement. This way of 
finding an optimal policy is called policy iteration.

\subsection{Value Iteration}
\noindent One drawback to policy iteration is that each of its iterations 
involves policy evaluation, which itself is a protracted iterative computation
requiring multiple sweeps through the state set.
\vspace{3mm}\newline We now truncate the policy evaluation such that it is
stopped after just one sweep~\cite{Sutton1998} (one update of each state):
\begin{align*}
    v_{k+1}(s)&=\underset{a}{\text{max}}\:q(s, a)\\
              &=\underset{a}{\text{max}}\:\mathbb{E}\left[
          R_{t+1}+\gamma v_k(S_{t+1})\;\middle|\;S_t=s, A_t=a\right]
\end{align*}
We keep iterate until \(v\) converges (to \(v_*\)), then we record the last
policy \(\pi(s) = \underset{a}{\text{argmax}}\:q(s, a)\).

\subsection{Generalized Policy Iteration~\cite{Sutton1998}}
\noindent As long as both processes, policy evaluation and policy improvement,
continue to update all states, the ultimate result is typically the 
same --- converge to the optimal value function and an optimal policy.
\vspace{3mm}\newline Generalized policy iteration (GPI) is used to refer to the
general idea of letting policy evaluation and policy improvement interact,
independent of the granularity of the two processes (instead of letting each
complete before the other begins).

@@

\subsection{Policy Gradient Methods~\cite{Sutton1998}}
\noindent When the state space is arbitrarily large, we cannot find an optimal 
policy or the optimal value function given the limited resources and time. 
Instead of keeping track of the value and the action(s) to select for each 
state (tabular methods), we can use parameterized functions, for example, 
artificial neural networks (ANNs), to approximate the value function and the 
policy function.
\vspace{3mm}\newline Previously methods learned the values of actions and then
select actions based on their estimated action values (q function). Now we
consider methods that instead learn a parameterized policy that select actions
without consulting a value function. A value function may still be used to
learn the policy parameter, but is not required for action selection.
\vspace{3mm}\newline We use \(\bm{\theta}\) and \(\bm{w}\) to denotes the
policy's parameter vector and value function's weight vector, respectively.
Then we have our policy \(\pi\left(a\;\middle|\;s, \bm{\theta}\right)
=\text{Pr}\left\{A_t=a\;\middle|\;S_t=s, \bm{\theta}_t=\bm{\theta}\right\}\) and
value function \(v_\pi(s, \bm{w})\). We also define the scalar performance
measure (objective) to be
\[J(\bm{\theta})=\begin{cases}
    \ v_{\pi_{\bm{\theta}}}(s_0); &\text{episodic case}\\
    \ \lim_{t \to \infty} \mathbb{E}
    \left[R_t\;\middle|\;S_0,A_{0:t-1}\sim\pi\right];
                                  & \text{continuing case}
\end{cases}\]
Policy gradient methods seeks to maximize this performance measure, so their
updates approximate gradient ascent in \(J\):
\[\bm{\theta}_{t+1}=\bm{\theta}_t+\alpha\widehat{\nabla J(\bm{\theta}_t)}\]
where \(\widehat{\nabla J(\bm{\theta}_t)}\) is a stochastic estimate whose
expection approximates the gradient of the performance measure w.r.t
\(\bm{\theta}_t\).

@@

\subsection{Training Result}
\noindent Figure~\ref{fig:a2c-out} shows the training result of 
Algorithm~\ref{algorithm:actor-critic} (300 moving average reward vs 
iteration), with 600 iterations (5 episodes per iteration), \(T=300\), 
\(\gamma=0.99\), \(\alpha_\theta=\alpha_{\bm{w}}=\num{3e-4}\).
\begin{figure*}
    \centering\includegraphics[scale=0.7]{a2c_out}
    \caption{Training Result for Algorithm a2c}
    \label{fig:a2c-out}
\end{figure*}
The graph shows that the moving average reward converges within only 50
iterations, that is, 250 episode, 75000 parameter updates for each network,
and stays quite stable afterwards.

\section{Actor-Critic, Reinforce, and PPO}
\subsection{Algorithms}
\noindent To further show how Actor-Critic algorithm performs, we compare it 
to another two popular algorithms: REINFORCE
(Algorithm~\ref{algorithm:reinforce}) and PPO (Algorithm~\ref{algorithm:ppo}).
The policy function of the two algorithms is implemented the same way as we 
did in Actor-Critic algorithm, and the same is true for the value function 
used in PPO method.
\begin{algorithm}
    \caption{Reinforce for financial problem~\cite{Sutton1998}}
    \label{algorithm:reinforce}
    \begin{algorithmic}[1]
        \State Initialize ANN policy parameterization
        \(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
        \State Step sizes \(\alpha_{\bm{\theta}}>0\)
        \Loop \hspace{0.3mm} (for each episode)
        \State Reset the environment
        \State Generate an episode 
        \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
        following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
        \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
        \EndFor
        \EndLoop
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}
    \caption{PPO for financial problem~\cite{SpinningUp2018}}
    \label{algorithm:ppo}
    \begin{algorithmic}[1]
        \State Initialize ANN policy parameterization
        \(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
        \State Initialize ANN state-value function parameterization
        \(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
        \State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
        \For{iteration \(k=0, 1, 2, ...\)}
        \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\),
        following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}_k\right)\), where
        \(\tau = S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\)
        \For{\(t=0, 1, ..., T\) in each trajectory \(\tau\) in 
        \(\mathcal{D}_k\)}
        \State \(G_t \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(A_t \gets G_t - \hat{v}(S_t)\)
        \EndFor
        \Loop \hspace{0.3mm} (update multiple times)\vspace{1mm}
        \State \(r \gets \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
        {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)}\)\vspace{1.2mm}
        \State \(L \gets \min{\left(
                rA_t,
                \text{clip}\left(r, 1-\epsilon, 1+\epsilon\right)A_t 
        \right)}\)\vspace{1mm}
        \State \(\bm{\theta} \gets \underset{\bm{\theta}}{\text{arg\:max}}
        \frac{1}{|\mathcal{D}_k|T}
        \sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} L\)
        \State \(\bm{w} \gets \underset{\bm{w}}{\text{arg\:min}}
        \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T}
        \Bigl(\hat{v}(S_t, \bm{w})-G_t\Bigr)^2\)
        \EndLoop
        \EndFor
    \end{algorithmic}
\end{algorithm}

\subsection{Training result}
\noindent Figure~\ref{fig:3-algorithms-out} shows the training 
result of these three algorithms (300 moving average reward vs iteration), 
with 600 iterations (5 episodes per iteration), \(T=300\), \(\gamma=0.99\),
\(\alpha_\theta=\num{3e-4}\), \(\alpha_{\bm{w}}=\num{3e-10}\) (if applicable),
\(\epsilon=0.5\) (if applicable).
\begin{figure*}
    \centering\includegraphics[scale=0.7]{out}
    \caption{Training Result Comparison for Three Algorithm}
    \label{fig:3-algorithms-out}
\end{figure*}
As we can see, the Actor-Critic algorithm converges the fastest, within 50 
iterations, followed by REINFORCE algorithm, which converges within 250 
iterations, despite a small drop of moving average around iteration 300, 
which is quickly corrected by the algorithm within 50 iterations. PPO 
algorithm converges the slowest, with about 300 iterations.

@@
